A weighted sampling algorithm for the design of RNA sequences with targeted secondary structure and nucleotides distribution

### 2013-03-11 ISMB/main_ISMB.tex

\documentclass{bioinfo}
\copyrightyear{2013}
\pubyear{2013}
%\usepackage[top=1in, bottom=1in,right=1in,left=1in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
%\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm} %pour les newtheorem
\usepackage{epstopdf}
\usepackage{xspace}
\usepackage{tcolorbox}
\usepackage{relsize}
\usepackage{natbib}
\usepackage[colorinlistoftodos]{todonotes}

@@

\usetikzlibrary{positioning}
\usetikzlibrary{matrix}
\usetikzlibrary{calc}
\usetikzlibrary{shapes}
\usetikzlibrary{arrows}
\usetikzlibrary{fit}
\usetikzlibrary{decorations}
\usetikzlibrary{snakes}
\pgfdeclarelayer{background}
\pgfdeclarelayer{backbackground}
\pgfdeclarelayer{foreground}
\pgfsetlayers{backbackground,background,main,foreground}

@@

\maketitle
%\section{}
%\subsection{}
\begin{abstract}
\section{Motivations:} The design of RNA sequences folding into predefined secondary structures is a milestone for many synthetic biology and gene therapy studies. Most of the current software uses similar local search strategies (i.e. a random seed is progressively adapted to acquire the desired folding propertues) and more importantly do not allow the user to control explicitly the nucleotide distribution such as the \GCContent in their sequences. However, the latter is an important criterion for large-scale applications as it could presumably be used to design sequences with better transcription rates and/or structural plasticity.

>>>>

\documentclass{bioinfo}
\copyrightyear{2013}
\pubyear{2013}
%\usepackage[top=1in, bottom=1in,right=1in,left=1in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
%\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm} %pour les newtheorem
\usepackage{epstopdf}
\usepackage{xspace}
\usepackage{tcolorbox}
\usepackage{relsize}
%\usepackage{natbib}
\usepackage[colorinlistoftodos]{todonotes}

@@

\usetikzlibrary{positioning}
\usetikzlibrary{matrix}
\usetikzlibrary{calc}
\usetikzlibrary{shapes}
\usetikzlibrary{arrows}
\usetikzlibrary{fit}
\usetikzlibrary{decorations}
\usetikzlibrary{snakes}
\usetikzlibrary{shadows}
\pgfdeclarelayer{background}
\pgfdeclarelayer{backbackground}
\pgfdeclarelayer{foreground}
\pgfsetlayers{backbackground,background,main,foreground}

@@

\maketitle
%\section{}
%\subsection{}
\begin{abstract}
\section{Motivations:} The design of RNA sequences folding into predefined secondary structures is a milestone for many synthetic biology and gene therapy studies. Most of the current software uses similar local search strategies (i.e. a random seed is progressively adapted to acquire the desired folding properties) and more importantly do not allow the user to control explicitly the nucleotide distribution such as the \GCContent in their sequences. However, the latter is an important criterion for large-scale applications as it could presumably be used to design sequences with better transcription rates and/or structural plasticity.

### 2013-03-11 ISMB/results_ISMB.tex
histo

\subsection{Implementation}
\label{sec:implementation}
Our software, \ourprog, is available at \oururl and was implemented in {\tt Python~2.7}. We used
\RNAinverse from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.

All time benchmarks were run on a single AMD Opteron(tm) 6278 Processor  at 2.4 GHz with cache of 512 KB.
The penalty $\beta$, associated with invalid base-pairs, was set to $15$.

@@

\subsection{Dataset}
To evaluate the quality of our method, we used secondary
structures from the \RNASTRAND database~\cite{andronescu2008rna}.
Those are known secondary structure from a variety of organisms.
We considered a subset of $50$ structures selected by Levin~\emph{et al}~\cite{Levin:2012kx}, 
whose length ranges between $20$ and $100$ nucleotides. 
 To ease the visualization of results, we clustered together structures
 having similar length, stacks density and proportion of free nucleotides in loops, leading to distributions of structures shown in Figure~\ref{fig:bins}.

>>>>

\subsection{Implementation}
\label{sec:implementation}
Our software, \ourprog, was implemented in {\tt Python~2.7}. We used
\RNAinverse from the \textit{Vienna Package 2.0}~\cite{Hofacker:1994}.
All time benchmarks were run on a single AMD Opteron(tm) 6278 Processor  at 2.4 GHz with cache of 512 KB.
The penalty $\beta$, associated with invalid base-pairs, was set to $15$.

@@

\subsection{Dataset}
To evaluate the quality of our method, we used secondary
structures from the \RNASTRAND database~\cite{andronescu2008rna}.
Those are known secondary structure from a variety of organisms.
We considered a subset of $50$ structures selected by~\cite{Levin:2012kx}, 
whose length ranges between $20$ and $100$ nucleotides. 
 To ease the visualization of results, we clustered together structures
 having similar length, stacks density and proportion of free nucleotides in loops, leading to distributions of structures shown in Figure~\ref{fig:bins}.
